---
title: "Broadband access in Texas"
output:
  html_document:
    df_print: paged
---

## Goals

- Introduce the official U.S. Census Bureau site data.census.gov.
- Explain API keys and set them up for census data.
- Introduce and explore the tidycensus package to pull census data into R.
- Introduce some mapping techniques using tidycensus. Introduce the concept of joining data together.
- Note some other packages and methods to use census data.

## Setup

```{r setup}
library(tidyverse)
library(tidycensus)
library(scales)
```

## Download variable lists

```{r vars}
v19_subject <- load_variables(2019, "acs5/subject", cache = TRUE)
```

## Test variable

Retrieved this from [table](https://data.census.gov/cedsci/table?q=broadband%20access&tid=ACSST5Y2019.S2801). Overall value for U.S.

Getting the broadband percent data

broadband of any type: S2801_C02_014

```{r test-var}
get_acs(
  year = 2019,
  variables = c(broadband = "S2801_C02_014"),
  geography = "us"
)
```

### Getting broadband for Texas counties

```{r tx-broadband}
broadband_tx <- get_acs(
  year = 2019,
  variables = c(broadband = "S2801_C02_014"),
  geography = "county",
  state = "TX"
)

broadband_tx
```

## Arrange to get counties with lowest/highest percentages of broadband

### Counties with lowest access

```{r lowest}
broadband_tx %>% 
  arrange(estimate) %>%
  head(10)
```

### Counties with highest access

```{r highest}
broadband_tx %>% 
  arrange(estimate %>% desc()) %>% 
  head(10)
```

## Get map data

```{r bb-geo}
broadband_tx_geo <- get_acs(
  year = 2019,
  variables = c(broadband = "S2801_C02_014"),
  geography = "county",
  state = "TX",
  geometry = TRUE
)

broadband_tx_geo
```

## Map broadband access in Texas

```{r bb-map}
broadband_tx_geo %>% 
  ggplot() +
  geom_sf(aes(fill = estimate), color = "white", size = .1) +
  theme_void() +
  labs(
    title = "Broadband access in Texas counties",
    caption = "Source: Census Bureau/ACS 5-year 2019"
  ) +
  scale_fill_distiller(
    palette = "Oranges",
    direction = 1,
    name = "% households\nwith broadband"
  ) +
  theme(plot.margin = unit(c(10,0,10,0), "pt"))
```
